---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
We start by unzipping the data from the GitHub repo, importing as a data frame,
and then populating a few vectors that will be useful for us later in the analysis.
```{r}
unzip('~/github/RepData_PeerAssessment1/activity.zip')
activityDF <- as.data.frame(read.table('activity.csv',
                                       header = T,
                                       sep = ",",
                                       colClasses = c("integer",
                                                      "Date",
                                                      "integer")
                                       )
                            )
uniqueDays <- as.Date(unique(activityDF$date))
intervals <- unique(activityDF$interval)
daysOfWeek <- weekdays(uniqueDays)
```


## What is mean total number of steps taken per day?
The mean total number of steps per day is 10766.19.
The median total number of steps per day is 10765.
```{r}
dayTotals <- numeric()
for (day in uniqueDays) {
      tempDate <- activityDF[activityDF$date == as.Date(day, origin="1970-01-01"),]
      daySum <- sum(tempDate$steps)
      dayTotals <- append(dayTotals, daySum)
}
dayMean <- mean(dayTotals, na.rm = T)
dayMedian <- median(dayTotals, na.rm = T)
print(dayMean)
```


## What is the average daily activity pattern?
```{r}
colMax <- function(data) sapply(data, max, na.rm = T)
dailyPattern <- data.frame(intervals=integer(),
                           averages=numeric())
for (i in intervals){
      tempInterval <- activityDF[activityDF$interval == i,]
      intervalAverage = mean(tempInterval$steps, na.rm = T)
      dailyPattern <- rbind(dailyPattern,data.frame(intervals = i, averages = intervalAverage))
}
plot(dailyPattern$intervals,
     dailyPattern$averages,
     type = "l",
     xlim = range(dailyPattern$intervals)
     )
maximum <- colMax(dailyPattern)
```


## Imputing missing values
```{r}
#activityDF$steps[is.na(activityDF$steps)] = mean(activityDF$steps, na.rm = T)

```


## Are there differences in activity patterns between weekdays and weekends?
```{r}

```
